\documentclass{article}
\pagestyle{empty}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{multicol}
\setlength{\oddsidemargin}{0in} \setlength{\evensidemargin}{0in}
\setlength{\topmargin}{0in} \setlength{\textheight}{8.5in}
\setlength{\textwidth}{6.5in}

\makeatletter
\renewcommand*\env@matrix[1][*\c@MaxMatrixCols c]{%
	\hskip -\arraycolsep
	\let\@ifnextchar\new@ifnextchar
	\array{#1}}
\makeatother

\begin{document}
\begin{flushleft}
	\bfseries{MATH 260, Linear Systems and Matrices, Fall `14}\\
	\bfseries{Activity 12:  Diagonalization}\\
%	\bfseries{Honor Code:} \hspace{3.5in}\bfseries{Names:}\\
\end{flushleft}
\begin{flushleft}

\section*{Warm-up:  Eigenvectors and Spaces of Repeated Roots}
%This is the matrix from Sect. 5.4, pg 339, #43

$\textbf{G}
=\begin{bmatrix}
1 & 0 & 0\\
-4 & 3 & 0 \\
-4 & 2 & 1
\end{bmatrix}
$
\hspace{0.5in}
The eigenvalues of \textbf{G} are $\lambda_{1,2}=1$ and $\lambda_3 = 3$. 

\vspace{0.25in}
1) If I had asked you to find the eigenvalues of \textbf{G}, you would NOT need to calculate $\det(\textbf{G}-\lambda \textbf{I})=0$. Explain why.

\vspace{1in}

2) Find the three eigenvectors associated with the given eigenvalues.

\vspace{3in}

3) What is the dimension of each corresponding eigenspace (recall, these are the geometric multiplicities of each eigenvalue)? What is the sum of the eigenspace dimensions for \textbf{G}?

\vspace{1in}

\section*{Diagonalized Equations}
Define two matrices: $\textbf{P} = \begin{bmatrix}
| & | & & |\\
\vec{v}_1 & \vec{v}_2 & \ldots &\vec{v}_n \\
| & | & & |
\end{bmatrix} $
\hspace{0.5in}
$\textbf{D} = \begin{bmatrix}
\lambda_1 & 0 & \ldots & 0 \\
0 & \lambda_2 & \ldots & 0 \\
\vdots & \vdots & \ddots & \vdots \\
0 & 0 & \ldots & \lambda_n 
\end{bmatrix} $

\vspace{0.1in}

We are going to check some matrix equations using \textbf{P}, $\textbf{P}^{-1}$ and \textbf{D}. 
For each of the following be sure to show some intermediate steps, since we already know what the final result should be!

\vspace{0.1in}

For \# 4-8 use the matrix: $\textbf{A} = \begin{bmatrix} 1 & 1 \\ 4 & 1 \end{bmatrix}$ \hspace{0.2in} with eigenvalues $\lambda_1 = 3$ and $\lambda_2 = -1$ with corresponding eigenvectors $\vec{v}_1 = \begin{bmatrix} 1 \\ 2 \end{bmatrix}$ and $\vec{v}_2 = \begin{bmatrix} 1 \\ -2 \end{bmatrix}$ respectively.

\vspace{0.2in}

4) Find $\textbf{AP}$ and $\textbf{PD}$. How do they compare?

\vspace{2.5in}

5) Find $\textbf{PDP}^{-1}$. What does it equal that we have already defined? \\
\textit{Hint: You already calculated $\textbf{PD}$ you just need to find $\textbf{P}^{-1}$, and do one multiplication}

\vspace{2.5in}

6) Find $\textbf{P}^{-1} \textbf{AP}$. What does it equal that we have already defined?\\
\textit{Hint: You only need to do ONE new multiplication here!}

\vspace{2in}

7) Write all three of the equations you have calculated above here, using the variables \textbf{A}, \textbf{P}, \textbf{D}, and $\textbf{P}^{-1}$.

\vspace{1in}

8)a) Define the matrix $\textbf{F}=\begin{bmatrix} \lambda_2 & 0 \\ 0 & \lambda_1 \end{bmatrix} = \begin{bmatrix} -1 & 0 \\ 0 & 3 \end{bmatrix}$, and use the same \textbf{P} as above. Find $\textbf{PFP}^{-1}$. 

\vspace{2.5in}

b) Does it matter to the equations you wrote in Problem 7 what order the eigenvalues and eigenvectors are in \textbf{P} and \textbf{D}? What relationship is required between \textbf{P} and \textbf{D}?

\vspace{1in}

\section*{Diagonalizing with Repeated Eigenvalues}
9) In the warm-up you found the eigenvectors for a matrix with a repeated eigenvalue, but which had 3 linearly independent eigenvectors. Pick your favorite equation from \# 6. Is the equation valid for \textbf{G}?

\vspace{2.5in}

%The matrix below is from Sect 5.4, pg 339, #29
10) The matrix $\begin{bmatrix} 3 & 1 \\ -1 & 5 \end{bmatrix}$ has a repeated eigenvalue of 4, with one eigenvector of $\vec{v}_1=\begin{bmatrix} 1 \\ 1 \end{bmatrix}$. \\
If we let $\textbf{P}=\begin{bmatrix} 1 & 1 \\ 1 & 1 \end{bmatrix}$ and $\textbf{D} = \begin{bmatrix} 4 & 0 \\ 0 & 4 \end{bmatrix}$. Does the equation $\textbf{A} = \textbf{PDP}^{-1}$ still hold? \\

\textit{Hint: Think about \textbf{P}. What is $det(\textbf{P})$? What does that tell you about $\textbf{P}^{-1}$? }

\vspace{2in}

11) If the results from \# 9 and \# 10 are generalizable (they are), what must be true about a matrix for us to diagonalize it?

\vspace{1in}

\end{flushleft}
\end{document}