\documentclass{article}
\usepackage{amsmath,amssymb,amsfonts,graphicx,nopageno}
\usepackage{multicol}
\usepackage{color}
\usepackage{soul}
\setlength{\oddsidemargin}{0in} \setlength{\evensidemargin}{0in}
\setlength{\topmargin}{0in} \setlength{\textheight}{8.5in}
\setlength{\textwidth}{6.5in}

\definecolor{Blue}{rgb}{0.3,0.3,0.9}

\begin{document}
\begin{flushleft}
	\bfseries{MATH 260, Final Test, Fall `14}\\
	\bfseries{November 20, 2014}\\
	\vspace{0.5in}
	\bfseries{Name:}\\
	\vspace{0.2in}
	\bfseries{Section:}\\
	\vspace{.5in}
		
	\bfseries{READ THESE DIRECTIONS:}
	\begin{itemize}
		\item{There are 6 problems, 5 total pages, 100 points on this test.}
		\item{Write your answers on this copy of the exam.   If you need more space, use the back of these sheets but clearly identify what problem it refers to.}
		\item{You must show your work.  Unjustified answers may not receive full credit.}
		\item{Even if you can't finish a problem, write down what you can!  You may receive partial credit.}
		\item{When you've finished the test, bring it to me.} %(in my office).}
		\item{Authorized aid:  Your brain.}
		\item{Un-authorized aid:  Anything not listed as authorized aid.  This includes calculators, notes, books, cell phones (FOR ANY REASON), etc.}
	\end{itemize}

\vspace{1in}

\bfseries{Honor Code:}

\pagebreak

1) (10 pts) Which of the following subsets of $\mathbb{R}^3$ are subspaces?  Justify your answer.

\vspace{.25in}

a) $S_1 = \{(x,y,z) \,|\, z = 2\}$.

\vspace{1.5in}

b) $S_2 = \{(x,y,z) \,|\, y = x + z\}$.

\vspace{1.5in}

2) (15 pts) State whether the following sets of vectors are linearly dependent or independent.  Justify your conclusion.

\vspace{.25in}

a) $\left\{ \begin{bmatrix}
0 \\ 1 \\ 0
\end{bmatrix}, \begin{bmatrix}
1 \\ 1 \\ 0
\end{bmatrix}, \begin{bmatrix}
2 \\ 0 \\ 1
\end{bmatrix}, \begin{bmatrix}
3 \\ 0 \\ 0
\end{bmatrix} \right\}$

\vspace{1in}

b) $\left\{ \begin{bmatrix}
1 \\ 0 \\ 1
\end{bmatrix}, \begin{bmatrix}
0 \\ 1 \\ 0
\end{bmatrix}, \begin{bmatrix}
1 \\ 1 \\ 0
\end{bmatrix} \right\}$

\vspace{1in}

c) $\left\{ \begin{bmatrix}
1 \\ 0 \\ 0 \\ 1 
\end{bmatrix}, \begin{bmatrix}
0 \\ 2 \\ 1 \\ 0
\end{bmatrix}, \begin{bmatrix}
0 \\ 1 \\ 1 \\ 0
\end{bmatrix}, \begin{bmatrix}
0 \\ 0 \\ 1 \\ 0
\end{bmatrix} \right\}$

\pagebreak

3) (10 pts) Find a basis for $S = \left\{\begin{bmatrix} a \\b \\b \end{bmatrix}  \,|\, a,b \in \mathbb{R}\right\}$ and give the dimension of $S$.

\vspace{2in}

4) Let $\textbf{A} = \begin{bmatrix}
1 & 1 & 1 \\
0 & 3 & 1 \\
0 & 0 & 1 \\
\end{bmatrix}$.

\vspace{.25in}

a) (5 pts) Find the eigenvalues of $\textbf{A}$.

\vspace{1.5in}

b) (10 pts) Find a basis for each eigenspace of $\textbf{A}$.

\vspace{3in}

c) (5 pts) Is $\textbf{A}$ diagonalizable?  Justify.

\pagebreak

5) (25 pts) Let $\textbf{A} = \begin{bmatrix}
1 & 1 & 1\\
1 & 3 & 7 \\
2 & 2 & 2
\end{bmatrix}$.  Find...

\vspace{.25in}
a) a basis for the null space of $\textbf{A}$.

\vspace{1.5in}
b) a basis for the column space of $\textbf{A}$.

\vspace{1.5in}

c) a basis for the column space of $\textbf{A}^T$.

\vspace{1.5in}

d) the rank($\textbf{A}$) and nullity($\textbf{A}$).

\vspace{1.5in}

e) Is $\begin{bmatrix}
2 \\
8 \\
4 
\end{bmatrix}$ in the column space of $\textbf{A}$?  Justify.

\pagebreak

6) (20 pts) Let $\textbf{A} = \begin{bmatrix}
2 & 3 \\
-1 & -2
\end{bmatrix}$.  Find an invertible matrix $\textbf{P}$ and a diagonal matrix $\textbf{D}$ such that $\textbf{A}$ = $\textbf{PDP}^{-1}$.  (NOTE:  You do NOT need to find $\textbf{P}^{-1}$ unless you want to check your work.)

\end{flushleft}
\end{document}