\documentclass{article}
\usepackage{amsmath,amssymb,amsfonts,graphicx,nopageno}
\usepackage{multicol}
\usepackage{color}
\usepackage{soul}
\setlength{\oddsidemargin}{0in} \setlength{\evensidemargin}{0in}
\setlength{\topmargin}{0in} \setlength{\textheight}{8.5in}
\setlength{\textwidth}{6.5in}

\definecolor{Blue}{rgb}{0.3,0.3,0.9}

\begin{document}
\begin{flushleft}
	\bfseries{MATH 260, Midterm Test, Fall `14}\\
	\bfseries{October 6, 2014}\\
	\vspace{0.5in}
	\bfseries{Name:}\\
	\vspace{0.2in}
	\bfseries{Section:}\\
	\vspace{.5in}
	
	\bfseries{READ THESE DIRECTIONS:}
	\begin{itemize}
		\item{There are 6 problems, 5 total pages, 100 points on this test.}
		\item{Write your answers on this copy of the exam.   If you need more space, use the back of these sheets but clearly identify what problem it refers to.}
		\item{You must show your work.  Unjustified answers may not receive full credit.}
		\item{Even if you can't finish a problem, write down what you can!  You may receive partial credit.}
		\item{When you've finished the test, bring it to me (in my office).}
		\item{Authorized aid:  Your brain.}
		\item{Un-authorized aid:  Anything not listed as authorized aid.  This includes calculators, notes, books, cell phones (FOR ANY REASON), etc.}
	\end{itemize}

\vspace{1in}

\bfseries{Honor Code:}

\pagebreak

1) (20 pts) Give an example of...

\vspace{0.2in}

a) a $3 \times 3$ matrix with rank 2.

\vspace{1in}

b) a $2 \times 2$ matrix with determinant 6.

\vspace{1in}

c) a $2 \times 2$ matrix that is singular.

\vspace{1in}

d) a linear system with infinitely many solutions (NOT its augmented matrix).

\vspace{1in}

2) (10 pts) Consider the linear system whose augmented matrix is of the form

\begin{equation*}
\begin{bmatrix}
1 & 2 & | & 3\\
0 & a & | & 1
\end{bmatrix}.
\end{equation*}

For what values of $a$, if any, will the system...

\vspace{.2in}

a) have a unique solution?

\vspace{0.75in}

b) have infinitely many solutions?

\vspace{0.75in}

c) be inconsistent?

\pagebreak

3) (20 pts) For the following linear system, give the augmented matrix and solve the system using Gauss-Jordan elimination (that is, get it into RREF first).

\begin{equation*}
\begin{array}{ccccccc}
x_{1} & + & x_{2} &  &  & = & 1\\
 &  & 3x_{2} & + & 3x_{3} & = & 6\\
-x_{1} & + & 3x_{2} & + & 5x_{3} & = & 7
\end{array}
\end{equation*}

\pagebreak

4) (15 pts) Evaluate the following if the operation is defined.  If it is not, say why.

\begin{equation*}
A = 
\begin{bmatrix}
2 & 1 \\
0 & -2 \\
1 & 0 
\end{bmatrix}, B = 
\begin{bmatrix}
-1 & 0 & 1\\
2 & 3 & 1
\end{bmatrix}, C =
\begin{bmatrix}
1 & 2 \\
-1 & 3
\end{bmatrix}.
\end{equation*}

a) $A + B^{T}$

\vspace{1in}

b) $BA$

\vspace{1in}

c) $BC$

\vspace{1in}

5) (15 pts) Find the inverse of the matrix if it is invertible.  Otherwise, show that the matrix is singular.

\vspace{0.2in}

$M = \begin{bmatrix}
1 & 2 & 0\\
-2 & -3 & 0\\
0 & 0 & \frac{1}{2}
\end{bmatrix}$

\pagebreak

6) Evaluate the following determinants.

\vspace{0.2in}

a) (8 pts) $\begin{vmatrix}
1 & 1 & 1\\
4 & 0 & 3\\
2 & 1 & 1
\end{vmatrix}$

\vspace{3in}

b) (12 pts) $\begin{vmatrix}
1 & 1 & 1 & 1\\
0 & 4 & 0 & 3\\
0 & 2 & 1 & 1\\
2 & 2 & 2 & 1
\end{vmatrix}$


\end{flushleft}
\end{document}